### 2024-12-25 download-football-data.qmd
20241225 - convert to calculate_stats() from nflfastR v5.0

# Downloaded Data - Raw
load(file = "./data/nfl_players_raw.RData")
load(file = "./data/nfl_teams_raw.RData")
load(file = "./data/nfl_rosters_raw.RData")
load(file = "./data/nfl_rosters_weekly_raw.RData")
load(file = "./data/nfl_schedules_raw.RData")
load(file = "./data/nfl_combine_raw.RData")
load(file = "./data/nfl_draftPicks_raw.RData")
load(file = "./data/nfl_depthCharts_raw.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
load(file = "./data/nfl_actualStats_weekly_raw.RData")
load(file = "./data/nfl_injuries_raw.RData")
load(file = "./data/nfl_snapCounts_raw.RData")
load(file = "./data/nfl_espnQBR_seasonal_raw.RData")
load(file = "./data/nfl_espnQBR_weekly_raw.RData")
load(file = "./data/nfl_nextGenStats_weekly_raw.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly_raw.RData")
load(file = "./data/nfl_playerContracts_raw.RData")
load(file = "./data/nfl_ftnCharting_raw.RData")
load(file = "./data/nfl_playerIDs_raw.RData")
load(file = "./data/nfl_rankings_draft_raw.RData")
load(file = "./data/nfl_rankings_weekly_raw.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp_raw.RData", fsep = ""))
load(file = "./data/players_projections_seasonal_raw.RData")

# Calculated Data - Raw
load(file = "./data/nfl_actualStats_career_raw.RData")
load(file = "./data/nfl_actualStats_seasonal_raw.RData")

# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
load(file = "./data/players_projectedPoints_seasonal.RData")

# Calculated Data - Processed
load(file = "./data/nfl_actualStats_career.RData")
load(file = "./data/nfl_actualStats_seasonal.RData")
```

@@

### Historical Weekly Actual Player Statistics {#sec-downloadActualStats}

@@

save(
  nfl_actualStats_offense_weekly_raw, nfl_actualStats_defense_weekly_raw, nfl_actualStats_kicking_weekly_raw,
  file = "./data/nfl_actualStats_weekly_raw.RData"
)
```

@@

@@

save(
  nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
  file = "./data/nfl_actualStats_weekly.RData"
)
```

@@

@@

Now, modify the scoring settings to match our league settings (based on [scoring settings](https://support.nfl.com/hc/en-us/articles/4989179237404-Scoring) for fantasy leagues on NFL.com):

@@

load("./data/nfl_actualStats_career_raw.RData")
```

@@

@@

load("./data/nfl_actualStats_seasonal_raw.RData")
```

@@

```{r}
#| eval: false
nfl_actualStats_weekly_raw <- calculate_stats(
  seasons = TRUE,
  summary_level = "week")

nfl_actualStats_weekly_defense_raw <- calculate_stats(
  seasons = TRUE,
  summary_level = "week",
  stat_type = "team")
```

@@

@@

#### Weekly {#sec-calculateActualPointsWeekly}

#### Seasonal {#sec-calculateActualPointsSeasonal}

@@

@@

```{r}
# Reshape from wide to long format
nfl_actualStats_offense_weekly_long <- nfl_actualStats_offense_weekly %>% 
  tidyr::pivot_longer(
    cols = c(team, opponent_team),
    names_to = "role",
    values_to = "team")

# Perform separate inner join operations for the home_team and away_team
nfl_actualStats_offense_weekly_home <- dplyr::inner_join(
  nfl_actualStats_offense_weekly_long,
  nfl_schedules,
  by = c("season","week","team" = "home_team")) %>% 
  mutate(home_away = "home_team")

nfl_actualStats_offense_weekly_away <- dplyr::inner_join(
  nfl_actualStats_offense_weekly_long,
  nfl_schedules,
  by = c("season","week","team" = "away_team")) %>% 
  mutate(home_away = "away_team")

nfl_actualStats_defense_weekly_home <- dplyr::inner_join(
  nfl_actualStats_defense_weekly,
  nfl_schedules,
  by = c("season","week","team" = "home_team")) %>% 
  mutate(home_away = "home_team")

nfl_actualStats_defense_weekly_away <- dplyr::inner_join(
  nfl_actualStats_defense_weekly,
  nfl_schedules,
  by = c("season","week","team" = "away_team")) %>% 
  mutate(home_away = "away_team")

nfl_actualStats_kicking_weekly_home <- dplyr::inner_join(
  nfl_actualStats_kicking_weekly,
  nfl_schedules,
  by = c("season","week","team" = "home_team")) %>% 
  mutate(home_away = "home_team")

nfl_actualStats_kicking_weekly_away <- dplyr::inner_join(
  nfl_actualStats_kicking_weekly,
  nfl_schedules,
  by = c("season","week","team" = "away_team")) %>% 
  mutate(home_away = "away_team")

# Combine the results of the join operations
nfl_actualStats_offense_weekly_schedules_long <- dplyr::bind_rows(
  nfl_actualStats_offense_weekly_home,
  nfl_actualStats_offense_weekly_away)

nfl_actualStats_defense_weekly_schedules_long <- dplyr::bind_rows(
  nfl_actualStats_defense_weekly_home,
  nfl_actualStats_defense_weekly_away)

nfl_actualStats_kicking_weekly_schedules_long <- dplyr::bind_rows(
  nfl_actualStats_kicking_weekly_home,
  nfl_actualStats_kicking_weekly_away)

# Reshape from long to wide
player_game_gameday_offense <- nfl_actualStats_offense_weekly_schedules_long %>%
  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
  tidyr::pivot_wider(
    names_from = home_away,
    values_from = team)

player_game_gameday_defense <- nfl_actualStats_defense_weekly_schedules_long %>%
  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
  tidyr::pivot_wider(
    names_from = home_away,
    values_from = team)

player_game_gameday_kicking <- nfl_actualStats_kicking_weekly_schedules_long %>%
  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
  tidyr::pivot_wider(
    names_from = home_away,
    values_from = team)

# Merge player birthdate and the game date
player_game_birthdate_gameday_offense <- dplyr::left_join(
  player_game_gameday_offense,
  unique(nfl_players[,c("gsis_id","birth_date")]),
  by = c("player_id" = "gsis_id")
)

player_game_birthdate_gameday_defense <- dplyr::left_join(
  player_game_gameday_defense,
  unique(nfl_players[,c("gsis_id","birth_date")]),
  by = c("player_id" = "gsis_id")
)

player_game_birthdate_gameday_kicking <- dplyr::left_join(
  player_game_gameday_kicking,
  unique(nfl_players[,c("gsis_id","birth_date")]),
  by = c("player_id" = "gsis_id")
)

player_game_birthdate_gameday_offense$birth_date <- lubridate::ymd(player_game_birthdate_gameday_offense$birth_date)
player_game_birthdate_gameday_offense$gameday <- lubridate::ymd(player_game_birthdate_gameday_offense$gameday)

player_game_birthdate_gameday_defense$birth_date <- lubridate::ymd(player_game_birthdate_gameday_defense$birth_date)
player_game_birthdate_gameday_defense$gameday <- lubridate::ymd(player_game_birthdate_gameday_defense$gameday)

player_game_birthdate_gameday_kicking$birth_date <- lubridate::ymd(player_game_birthdate_gameday_kicking$birth_date)
player_game_birthdate_gameday_kicking$gameday <- lubridate::ymd(player_game_birthdate_gameday_kicking$gameday)

# Calculate player's age for a given week as the difference between their birthdate and the game date
player_game_birthdate_gameday_offense$age <- lubridate::interval(
  start = player_game_birthdate_gameday_offense$birth_date,
  end = player_game_birthdate_gameday_offense$gameday
) %>% 
  lubridate::time_length(unit = "years")

player_game_birthdate_gameday_defense$age <- lubridate::interval(
  start = player_game_birthdate_gameday_defense$birth_date,
  end = player_game_birthdate_gameday_defense$gameday
) %>% 
  lubridate::time_length(unit = "years")

player_game_birthdate_gameday_kicking$age <- lubridate::interval(
  start = player_game_birthdate_gameday_kicking$birth_date,
  end = player_game_birthdate_gameday_kicking$gameday
) %>% 
  lubridate::time_length(unit = "years")

# Merge with Pro Football Reference Data on Player Age by Season
player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

# Set age as first non-missing value from calculation above or from PFR
player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)

player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)

player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)

>>>>

# Downloaded Data - Raw
load(file = "./data/nfl_players_raw.RData")
load(file = "./data/nfl_teams_raw.RData")
load(file = "./data/nfl_rosters_raw.RData")
load(file = "./data/nfl_rosters_weekly_raw.RData")
load(file = "./data/nfl_schedules_raw.RData")
load(file = "./data/nfl_combine_raw.RData")
load(file = "./data/nfl_draftPicks_raw.RData")
load(file = "./data/nfl_depthCharts_raw.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
load(file = "./data/nfl_actualStats_position_weekly_raw.RData")
load(file = "./data/nfl_actualStats_weekly_raw.RData")
load(file = "./data/nfl_injuries_raw.RData")
load(file = "./data/nfl_snapCounts_raw.RData")
load(file = "./data/nfl_espnQBR_seasonal_raw.RData")
load(file = "./data/nfl_espnQBR_weekly_raw.RData")
load(file = "./data/nfl_nextGenStats_weekly_raw.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly_raw.RData")
load(file = "./data/nfl_playerContracts_raw.RData")
load(file = "./data/nfl_ftnCharting_raw.RData")
load(file = "./data/nfl_playerIDs_raw.RData")
load(file = "./data/nfl_rankings_draft_raw.RData")
load(file = "./data/nfl_rankings_weekly_raw.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp_raw.RData", fsep = ""))
load(file = "./data/players_projections_seasonal_raw.RData")

# Calculated Data - Raw
load(file = "./data/nfl_actualStats_career_raw.RData")
load(file = "./data/nfl_actualStats_seasonal_raw.RData")
load(file = "./data/nfl_actualFantasyPoints_career_raw.RData")
load(file = "./data/nfl_actualFantasyPoints_seasonal_raw.RData")
load(file = "./data/nfl_actualFantasyPoints_weekly_raw.RData")

# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = "./data/nfl_actualStats_position_weekly.RData")
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
load(file = "./data/players_projectedPoints_seasonal.RData")

# Calculated Data - Processed
load(file = "./data/nfl_actualStats_career.RData")
load(file = "./data/nfl_actualStats_seasonal.RData")
load(file = "./data/nfl_actualFantasyPoints_career.RData")
load(file = "./data/nfl_actualFantasyPoints_seasonal.RData")
load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
```

@@

### Historical Actual Player Statistics {#sec-downloadActualStats}

#### Career Statistics {#sec-downloadActualCareerStats}

#### Season-by-Season Statistics {#sec-downloadActualSeasonalStats}

#### Week-by-Week Statistics {#sec-downloadActualWeeklyStats}

@@

save(
  nfl_actualStats_offense_weekly_raw, nfl_actualStats_defense_weekly_raw, nfl_actualStats_kicking_weekly_raw,
  file = "./data/nfl_actualStats_position_weekly_raw.RData"
)
```

@@

load(file = "./data/nfl_actualStats_position_weekly_raw.RData")
```

@@

save(
  nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
  file = "./data/nfl_actualStats_position_weekly.RData"
)
```

@@

load(file = "./data/nfl_actualStats_position_weekly.RData")
```

@@

Now, modify the scoring settings to match your league settings.
Below, we use the [scoring settings](https://support.nfl.com/hc/en-us/articles/4989179237404-Scoring) for fantasy leagues on NFL.com, which happen to be point-per-reception leagues (i.e., PPR leagues):

@@

load(file = "./data/nfl_actualStats_career_raw.RData")
```

@@

::: {#nte-downloadSeasonalActualStats .callout-note title="Downloading season-by-season statistics"}
Note: the following code takes a while to run.
:::

@@

load(file = "./data/nfl_actualStats_seasonal_raw.RData")
```

@@

::: {#nte-downloadWeeklyActualStats .callout-note title="Downloading week-by-week statistics"}
Note: the following code takes a while to run.
:::

```{r}
#| eval: false
nfl_actualStats_weekly_player_raw <- calculate_stats(
  seasons = TRUE,
  summary_level = "week")

nfl_actualStats_weekly_team_raw <- calculate_stats(
  seasons = TRUE,
  summary_level = "week",
  stat_type = "team")
```

@@

save(
  nfl_actualStats_weekly_player_raw, nfl_actualStats_weekly_team_raw,
  file = "./data/nfl_actualStats_weekly_raw.RData"
)
```

```{r}
nfl_actualStats_weekly_player <- nfl_actualStats_weekly_player_raw %>% 
  select(player_id, season, week, everything()) %>% 
  arrange(player_display_name, player_id, season, week)

nfl_actualStats_weekly_team <- nfl_actualStats_weekly_team_raw %>% 
  select(season, week, team, everything()) %>% 
  arrange(season, week, team)
```

```{r}
nfl_actualStats_weekly_player %>% 
  group_by(player_id, season, week) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualStats_weekly_team %>% 
  group_by(team, season, week) %>% 
  filter(n() > 1) %>% 
  head()
```

save(
  nfl_actualStats_weekly_player, nfl_actualStats_weekly_team,
  file = "./data/nfl_actualStats_weekly.RData"
)
```

@@

#### Week-by-Week {#sec-calculateActualPointsWeekly}

```{r}
nfl_actualFantasyPoints_player_weekly_seasonalList <- list()
nfl_actualFantasyPoints_dst_weekly_seasonalList <- list()
```

::: {#nte-calculateActualFantasyPointsWeekly .callout-note title="Calculating players' week-by-week fantasy points"}
Note: the following code takes a while to run.
:::

#nfl_actualFantasyPoints_weekly_raw <- ffanalytics:::actual_points_scoring(
#  season = 2023,
#  summary_level = c("week"),
#  stat_type = c("player", "dst", "team"), 
#  season_type = c("REG", "POST", "REG+POST"),
#  scoring_rules = scoring_obj,
#  vor_baseline = NULL,
#  rename_colums = FALSE
#)

for(i in 1:length(seasons)){
  # Compute actual statistics by season
  nfl_actualFantasyPoints_player_weekly_seasonalList[[i]] <- 
    ffanalytics:::actual_points_scoring(
      season = seasons[i],
      summary_level = c("week"),
      stat_type = c("player"), 
      #season_type = c("REG"),
      scoring_rules = scoring_obj,
      vor_baseline = NULL,
      rename_colums = FALSE)

  nfl_actualFantasyPoints_dst_weekly_seasonalList[[i]] <- 
    ffanalytics:::actual_points_scoring(
      season = seasons[i],
      summary_level = c("week"),
      stat_type = c("dst"), 
      #season_type = c("REG"),
      scoring_rules = scoring_obj,
      vor_baseline = NULL,
      rename_colums = FALSE)

  nfl_actualFantasyPoints_player_weekly_seasonalList[[i]]$season <- seasons[i]
  nfl_actualFantasyPoints_dst_weekly_seasonalList[[i]]$season <- seasons[i]

  print(
    paste("Completed computing actual fantasy points for season: ", seasons[i], sep = ""))

nfl_actualFantasyPoints_player_weekly_raw <- nfl_actualFantasyPoints_player_weekly_seasonalList %>% 
  dplyr::bind_rows()

nfl_actualFantasyPoints_dst_weekly_raw <- nfl_actualFantasyPoints_dst_weekly_seasonalList %>% 
  dplyr::bind_rows()
```

save(
  nfl_actualFantasyPoints_player_weekly_raw, nfl_actualFantasyPoints_dst_weekly_raw,
  file = "./data/nfl_actualFantasyPoints_weekly_raw.RData"
)
```

load(file = "./data/nfl_actualFantasyPoints_weekly_raw.RData")
```

The `nfl_actualFantasyPoints_weekly` objects are in `player`-`season`-`week` (or `team`-`season`-`week`) form.
That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week` (or `team`-`season`-`week`).
Let's rearrange the data accordingly:

```{r}
nfl_actualFantasyPoints_player_weekly <- nfl_actualFantasyPoints_player_weekly_raw %>% 
  rename(
    fantasyPoints = raw_points,
    team = recent_team) %>% 
  select(player_id, season, week, everything()) %>% 
  arrange(player_display_name, player_id, season, week)

nfl_actualFantasyPoints_dst_weekly <- nfl_actualFantasyPoints_dst_weekly_raw %>% 
  rename(
    fantasyPoints = raw_points,
    team = recent_team) %>% 
  select(season, week, team, everything()) %>% 
  arrange(season, week, team)
```

```{r}
nfl_actualFantasyPoints_player_weekly %>% 
  group_by(player_id, season, week) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualFantasyPoints_dst_weekly %>% 
  group_by(team, season, week) %>% 
  filter(n() > 1) %>% 
  head()
```

save(
  nfl_actualFantasyPoints_player_weekly, nfl_actualFantasyPoints_dst_weekly,
  file = "./data/nfl_actualFantasyPoints_weekly.RData"
)
```

load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
```

nfl_actualStats_offense_weekly
nfl_actualStats_defense_weekly
nfl_actualStats_kicking_weekly

test_weekly_offense <- ffanalytics:::actual_points_scoring(
  nflr_player_stats_df = nfl_actualStats_offense_weekly %>% filter(season == 2023),
  nflr_pbp_data = nfl_pbp %>% filter(season == 2023),
  scoring_rules = scoring_obj,
  vor_baseline = NULL,
  rename_colums = TRUE)

test_weekly_defense <- ffanalytics:::actual_points_scoring(
  nflr_player_stats_df = nfl_actualStats_defense_weekly %>% filter(season == 2023),
  nflr_pbp_data = nfl_pbp %>% filter(season == 2023),
  scoring_rules = scoring_obj,
  vor_baseline = NULL,
  rename_colums = TRUE)

test_weekly_kicking <- ffanalytics:::actual_points_scoring(
  nflr_player_stats_df = nfl_actualStats_kicking_weekly %>% filter(season == 2023),
  nflr_pbp_data = nfl_pbp %>% filter(season == 2023),
  scoring_rules = scoring_obj,
  vor_baseline = NULL,
  rename_colums = TRUE)
```

#### Season-by-Season {#sec-calculateActualPointsSeasonal}

```{r}
nfl_actualFantasyPoints_player_seasonal_seasonalList <- list()
nfl_actualFantasyPoints_dst_seasonal_seasonalList <- list()
```

::: {#nte-calculateActualFantasyPointsSeasonal .callout-note title="Calculating players' season-by-season fantasy points"}
Note: the following code takes a while to run.
:::

#nfl_actualFantasyPoints_seasonal_raw <- ffanalytics:::actual_points_scoring(
#  season = 2023,
#  summary_level = c("season"),
#  stat_type = c("player", "dst", "team"), 
#  season_type = c("REG"),
#  scoring_rules = scoring_obj,
#  vor_baseline = NULL,
#  rename_colums = TRUE
#)

for(i in 1:length(seasons)){
  # Compute actual statistics by season
  nfl_actualFantasyPoints_player_seasonal_seasonalList[[i]] <- 
    ffanalytics:::actual_points_scoring(
      season = seasons[i],
      summary_level = c("season"),
      stat_type = c("player"), 
      season_type = c("REG"),
      scoring_rules = scoring_obj,
      vor_baseline = NULL,
      rename_colums = FALSE)

  nfl_actualFantasyPoints_dst_seasonal_seasonalList[[i]] <- 
    ffanalytics:::actual_points_scoring(
      season = seasons[i],
      summary_level = c("season"),
      stat_type = c("dst"), 
      season_type = c("REG"),
      scoring_rules = scoring_obj,
      vor_baseline = NULL,
      rename_colums = FALSE)

  nfl_actualFantasyPoints_player_seasonal_seasonalList[[i]]$season <- seasons[i]
  nfl_actualFantasyPoints_dst_seasonal_seasonalList[[i]]$season <- seasons[i]

  print(
    paste("Completed computing actual fantasy points for season: ", seasons[i], sep = ""))

nfl_actualFantasyPoints_player_seasonal_raw <- nfl_actualFantasyPoints_player_seasonal_seasonalList %>% 
  dplyr::bind_rows()

nfl_actualFantasyPoints_dst_seasonal_raw <- nfl_actualFantasyPoints_dst_seasonal_seasonalList %>% 
  dplyr::bind_rows()
```

save(
  nfl_actualFantasyPoints_player_seasonal_raw, nfl_actualFantasyPoints_dst_seasonal_raw,
  file = "./data/nfl_actualFantasyPoints_seasonal_raw.RData"
)
```

load(file = "./data/nfl_actualFantasyPoints_seasonal_raw.RData")
```

The `nfl_actualFantasyPoints_seasonal` objects are in `player`-`season` (or `team`-`season`) form.
That is, each row should be uniquely identified by the combination of `player_id` and `season` (or `team`-`season`).
Let's rearrange the data accordingly:

```{r}
nfl_actualFantasyPoints_player_seasonal <- nfl_actualFantasyPoints_player_seasonal_raw %>% 
  rename(fantasyPoints = raw_points) %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)

nfl_actualFantasyPoints_dst_seasonal <- nfl_actualFantasyPoints_dst_seasonal_raw %>% 
  rename(
    fantasyPoints = raw_points,
    team = recent_team) %>% 
  select(season, team, everything()) %>% 
  arrange(season, team)
```

```{r}
nfl_actualFantasyPoints_player_seasonal %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualFantasyPoints_dst_seasonal %>% 
  group_by(team, season) %>% 
  filter(n() > 1) %>% 
  head()
```

save(
  nfl_actualFantasyPoints_player_seasonal, nfl_actualFantasyPoints_dst_seasonal,
  file = "./data/nfl_actualFantasyPoints_seasonal.RData"
)
```

load(file = "./data/nfl_actualFantasyPoints_seasonal.RData")
```

@@

```{r}
nfl_actualFantasyPoints_player_career_raw <- nfl_actualFantasyPoints_player_seasonal %>% 
  group_by(player_id) %>% 
  summarise(fantasyPoints = sum(fantasyPoints, na.rm = TRUE))

nfl_actualFantasyPoints_player_career_raw <- full_join(
  nfl_actualFantasyPoints_player_career_raw,
  nfl_players %>% select(gsis_id, display_name, first_name, last_name, position_group, position),
  by = c("player_id" = "gsis_id")
)
```

save(
  nfl_actualFantasyPoints_player_career,
  file = "./data/nfl_actualFantasyPoints_career_raw.RData"
)
```

load(file = "./data/nfl_actualFantasyPoints_career_raw.RData")
```

```{r}
nfl_actualFantasyPoints_player_career <- nfl_actualFantasyPoints_player_career_raw %>% 
  select(player_id, display_name, first_name, last_name, position_group, position, fantasyPoints)
```

save(
  nfl_actualFantasyPoints_player_career,
  file = "./data/nfl_actualFantasyPoints_career.RData"
)
```

load(file = "./data/nfl_actualFantasyPoints_career.RData")
```

@@

```{r}
# Reshape from wide to long format
nfl_actualFantasyPoints_player_weekly_long <- nfl_actualFantasyPoints_player_weekly %>% 
  tidyr::pivot_longer(
    cols = c(team, opponent_team),
    names_to = "role",
    values_to = "team")

# Perform separate inner join operations for the home_team and away_team
nfl_actualFantasyPoints_player_weekly_home <- dplyr::inner_join(
  nfl_actualFantasyPoints_player_weekly_long,
  nfl_schedules,
  by = c("season","week","team" = "home_team")) %>% 
  mutate(home_away = "home_team")

nfl_actualFantasyPoints_player_weekly_away <- dplyr::inner_join(
  nfl_actualFantasyPoints_player_weekly_long,
  nfl_schedules,
  by = c("season","week","team" = "away_team")) %>% 
  mutate(home_away = "away_team")

# Combine the results of the join operations
nfl_actualFantasyPoints_player_weekly_schedules_long <- dplyr::bind_rows(
  nfl_actualFantasyPoints_player_weekly_home,
  nfl_actualFantasyPoints_player_weekly_away)

# Reshape from long to wide
player_game_gameday <- nfl_actualFantasyPoints_player_weekly_schedules_long %>%
  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
  tidyr::pivot_wider(
    names_from = home_away,
    values_from = team)

# Merge player birthdate and the game date
player_game_birthdate_gameday <- dplyr::left_join(
  player_game_gameday,
  unique(nfl_players[,c("gsis_id","birth_date")]),
  by = c("player_id" = "gsis_id")
)

player_game_birthdate_gameday$birth_date <- lubridate::ymd(player_game_birthdate_gameday$birth_date)
player_game_birthdate_gameday$gameday <- lubridate::ymd(player_game_birthdate_gameday$gameday)

# Calculate player's age for a given week as the difference between their birthdate and the game date
player_game_birthdate_gameday$age <- lubridate::interval(
  start = player_game_birthdate_gameday$birth_date,
  end = player_game_birthdate_gameday$gameday
) %>% 
  lubridate::time_length(unit = "years")

# Merge with Pro Football Reference Data on Player Age by Season
player_game_birthdate_gameday <- player_game_birthdate_gameday %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

# Set age as first non-missing value from calculation above or from PFR
player_game_birthdate_gameday <- player_game_birthdate_gameday %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)

# Calculate ageCentered and ageCenteredQuadratic
player_game_birthdate_gameday$ageCentered20 <- player_game_birthdate_gameday$age - 20
player_game_birthdate_gameday$ageCentered20Quadratic <- player_game_birthdate_gameday$ageCentered20 ^ 2

# Merge with player info
player_age <- dplyr::left_join(
  player_game_birthdate_gameday,
  nfl_players %>% select(-birth_date, -team_abbr, - team_seq),
  by = c("player_id" = "gsis_id"))

# Add game_id to weekly stats to facilitate merging
nfl_actualFantasyPoints_player_weekly <- nfl_actualFantasyPoints_player_weekly %>% 
  dplyr::left_join(
    player_age[,c("season","week","player_id","game_id")],
    by = c("season","week","player_id"))

# Merge with player weekly stats
player_stats_weekly <- dplyr::full_join(
  player_age %>% select(-position, -position_group),
  nfl_actualFantasyPoints_player_weekly,
  by = c("season","week","player_id","game_id"))

player_stats_weekly$total_years_of_experience <- as.integer(player_stats_weekly$years_of_experience)

player_stats_weekly$years_of_experience <- NULL

distinct_seasons <- player_stats_weekly %>%
  dplyr::select(player_id, season) %>%
  dplyr::distinct() %>% 
  dplyr::left_join(
    nfl_players[,c("gsis_id","years_of_experience")],
    by = c("player_id" = "gsis_id")
  ) %>% 
  dplyr::mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
  dplyr::select(-years_of_experience)

years_of_experience <- distinct_seasons %>% 
  dplyr::arrange(player_id, -season) %>% 
  dplyr::group_by(player_id) %>%
  dplyr::mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
  dplyr::ungroup()

years_of_experience$years_of_experience[which(years_of_experience$years_of_experience < 0)] <- 0

player_stats_weekly <- player_stats_weekly %>% 
  dplyr::left_join(
    years_of_experience[,c("player_id","season","years_of_experience")],
    by = c("player_id","season")
  )
```
